A multiverse analysis of the associations between internet use and well-being

### 2023-09-12 ms.Rmd
New journal

---
title: 'A multiverse analysis of the associations between internet use and well-being'
shorttitle: Internet and well-being
leftheader: Internet and well-being
author: 
  - name: Matti Vuorre
    affiliation: "1,2"
    corresponding: yes
    address: Tilburg University
    email: m.j.vuorre@tilburguniversity.edu
  - name: Andrew K. Przybylski
    affiliation: "2"
    email: andy.przybylski@oii.ox.ac.uk
affiliation:
  - id: 1
    institution: Tilburg School of Social and Behavioral Sciences, Tilburg University
  - id: 2
    institution: Oxford Internet Institute, University of Oxford
authornote: |
  \noindent \textbf{This manuscript is not yet peer-reviewed.}

keywords: well-being, internet technology, technology effects
wordcount: '`r scales::number(wordcountaddin:::word_count("ms.Rmd"), big.mark = ",")`'
bibliography: references.bib
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: jou
csl: "pnas.csl"
header-includes:
  - |
    \setlength{\parskip}{0pt}
    \setlength\parindent{24pt}
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
    keep_tex: false
---

@@

---
abstract: |
  \paragraph{Abstract}
  Internet technologies' and platforms' potential psychological consequences remain debated. While these technologies have spurred new forms of commerce, education, and leisure, many are worried that they might negatively affect individuals by, for example, displacing time spent on other healthy activities. Relevant findings to date have been inconclusive and of limited geographic and demographic scope. We examined whether having (mobile) internet access or actively using the internet predicted eight well-being outcomes from `r yr` among `r number2(nrow(dat))` individuals across `r length(unique(dat$country))` countries. We first queried the extent to which well-being varied as a function of internet connectivity. Then, we examined these associations' robustness in a multiverse of `r number2(nrow(specs))` analysis specifications. `r percent(mvs$Positive[9], .1)` of these resulted in positive and statistically significant associations between internet connectivity and well-being. These results indicate that internet access and use predict well-being positively and independently from a set of plausible alternatives.
  \paragraph{Significance statement}
  Are internet technologies, such as smartphones and social media, psychologically harmful? Past results have offered mixed evidence and, critically, focused on a narrow slice of the global population. Stakeholders from international health bodies to technology companies are investing in regulation and tools to protect internet users' well-being. We examined 2.5 million individual's psychosocial well-being and internet adoption and use over a period of fifteen years. We find that (mobile) internet adoption and use predict positive well-being and that this association is robust across thousands of adjustments for alternative explanations and demographic subgroups. While merely suggestive of causal associations, our results highlight a need for more transparent access to large databases currently held behind the closed doors of the technology industry.
---

@@

# Open Practices Statement

@@

>>>>

---
title: 'A multiverse analysis of the associations between internet use and well-being'
shorttitle: Internet and well-being
leftheader: Internet and well-being
author: 
  - name: Matti Vuorre
    affiliation: "1,2"
    corresponding: yes
    address: Tilburg University
    email: m.j.vuorre@tilburguniversity.edu
  - name: Andrew K. Przybylski
    affiliation: "2"
    email: andy.przybylski@oii.ox.ac.uk
affiliation:
  - id: 1
    institution: Tilburg School of Social and Behavioral Sciences, Tilburg University
  - id: 2
    institution: Oxford Internet Institute, University of Oxford
authornote: |
  \noindent \textbf{This manuscript is not yet peer-reviewed. The preprint is available at https://psyarxiv.com/jp5nd/}
  We used the proprietary Gallup World Poll dataset, which is available to subscribing institutions through the Gallup website. All our code and a simulated GWP dataset is available at https://doi.org/10.5281/zenodo.7774923. We did not pre-specify the sample sizes, but instead used all the available data in the Gallup database (with the exclusions detailed below). This study was not pre-registered.
keywords: well-being, internet technology, technology effects
wordcount: '`r scales::number(wordcountaddin:::word_count("ms.Rmd"), big.mark = ",")`'
bibliography: references.bib
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: doc
header-includes:
  - |
    \setlength{\parskip}{0pt}
    \setlength\parindent{24pt}
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
    keep_tex: false
---

@@

---
abstract: |
  Internet technologies' and platforms' potential psychological consequences remain debated. While these technologies have spurred new forms of commerce, education, and leisure, many are worried that they might negatively affect individuals by, for example, displacing time spent on other healthy activities. Relevant findings to date have been inconclusive and of limited geographic and demographic scope. We examined whether having (mobile) internet access or actively using the internet predicted eight well-being outcomes from `r yr` among `r number2(nrow(dat))` individuals across `r length(unique(dat$country))` countries. We first queried the extent to which well-being varied as a function of internet connectivity. Then, we examined these associations' robustness in a multiverse of `r number2(nrow(specs))` analysis specifications. `r percent(mvs$Positive[9], .1)` of these resulted in positive and statistically significant associations between internet connectivity and well-being. These results indicate that internet access and use predict well-being positively and independently from a set of plausible alternatives.
---

@@

@@

# Data availability statement
